### 2014-04-04 Presentazione/Presentazione_SynCoP.tex
Added more on parameter settings

\documentclass{beamer}
\usetheme{ut}
\usepackage{pgf}
%\usepackage{slashbox}
\usepackage{array}
\usepackage{calc}
\usepackage{tikz}
\usetikzlibrary{positioning}

\pgfdeclareimage[width=.9\textwidth]{signaling-schematic}{images/signaling_schematic}
\pgfdeclareimage[width=.6\textwidth]{activity00}{images/protein_activity00}
\pgfdeclareimage[width=.6\textwidth]{activity01}{images/protein_activity01}
\pgfdeclareimage[width=.6\textwidth]{activity02}{images/protein_activity02}
\pgfdeclareimage[width=.6\textwidth]{activity03}{images/protein_activity03}
\pgfdeclareimage[width=.6\textwidth]{activity04}{images/protein_activity04}
\pgfdeclareimage[width=.9\textwidth]{ta}{images/ta}
\pgfdeclareimage[width=.8\textwidth]{case-study}{images/mapk_model_egf4}
\pgfdeclareimage[width=.9\textwidth]{grafico-ruvido}{images/10levels}
\pgfdeclareimage[width=.9\textwidth]{grafico-liscio}{images/100levels}
\pgfdeclareimage[width=.9\textwidth]{animo-logo}{images/animo_logo}
\pgfdeclareimage[width=.3\textwidth]{animo-qrcode}{images/animo_site_qrcode}
\pgfdeclareimage[width=.9\textwidth]{animo-user-interface}{images/animo_user_interface}
\pgfdeclareimage[width=.75\textwidth]{animo-workflow}{images/animo_simulation_workflow}
\pgfdeclareimage[width=2cm]{activity-legend}{images/activity_legend}
\pgfdeclareimage[width=.9\textwidth]{ta-reactant-centered}{images/Reactant-centered}

\newenvironment{changemargin}[2]{%
\begin{list}{}{%
\setlength{\topsep}{0pt}%
\setlength{\leftmargin}{#1}%
\setlength{\rightmargin}{#2}%
\setlength{\listparindent}{\parindent}%
\setlength{\itemindent}{\parindent}%
\setlength{\parsep}{\parskip}%
}%
\item[]}{\end{list}}

\newcommand{\titledFrameImage}[2]{
\begin{frame}{#1}
%\begin{changemargin}{-1cm}{-1cm}
\begin{center}
\includegraphics[width=108mm,height=\textheight,keepaspectratio]{#2}
\end{center}
%\end{changemargin}
\end{frame}
}

\newcommand{\plainFrameImage}[1]{
\begin{frame}[plain]
%\begin{changemargin}{-1cm}{-1cm}
\begin{center}
\includegraphics[width=108mm,height=76mm,keepaspectratio]{#1}
\end{center}
%\end{changemargin}
\end{frame}
}

\newcommand{\maxFrameImage}[1]{
    %\setbeamertemplate{background}[ut][]
    \begin{frame}[plain]
        \begin{changemargin}{-2cm}{-2cm}
        \begin{center}
            \includegraphics[width=\paperwidth,height=\paperheight,keepaspectratio]{#1}
        \end{center}
        \end{changemargin}
    \end{frame}
    %\setbeamertemplate{background}[ut][0]
}

\newcommand*{\xdash}[1][3em]{\rule[0.5ex]{#1}{0.55pt}}

\makeatletter
\setbeamertemplate{footline}
{%
 \hspace{\beamer@leftmargin}
\hspace{-4.6ex}
\includegraphics[height=4mm,page=1]{beamerthemeutresources}
\hspace*{2ex}
\hfill
\begin{beamercolorbox}[ht=2cm,wd=9cm,sep=0cm]{fg=black,bg=red}
\hfill
%  [ \insertshortauthor, \insertshortinstitute\ ]
\tiny\insertfootlinetext
\hspace{2ex}
%\tiny\insertframenumber/\inserttotalframenumber
\hspace*{\beamer@rightmargin}
\vspace*{1.5ex}
\end{beamercolorbox}
}
\makeatother

\title{Setting parameters for biological~models with ANIMO}
\author{\small \underline{Stefano~Schivo}, Jetse~Scholma, Marcel~Karperien, Janine~N.~Post, Jaco~van~de~Pol, Rom~Langerak}
\institute{University of Twente,\\ Enschede, The Netherlands}
\footlinetext{ANIMO \--\ Analysis of Networks with Interactive MOdeling}
\date[SynCoP2014]{SynCoP 2014}
%\date{}

% A che domande biologiche rispondiamo con ANIMO
% Approccio: interaction-based, discrete levels, sottolinea la concorrenza
% Magari evitare di mostrare modelli TA
% Come funziona la modellazione in ANIMO: topologia, scenari, parametri (con esempi) (tipo Gene?)
% Come funziona l'esperimento in silico (simulazione, model checking, grafici sim+data, slider)
% Demo breve
% Piani futuri: generazione di ipotesi, model checking (esempi??), high-performance analysis
%\listfiles

\begin{document}

\maketitle

\frame{ \frametitle{Signalling Pathways}
  \begin{center}
    %\pgfuseimage{pathway-semplificato}
    \begin{minipage}[c]{.9\textwidth}%
      \pgfuseimage{signaling-schematic}\\[-1ex]%
      {\tiny Credit: National Science Foundation}%
    \end{minipage}
  \end{center}
}

  \maxFrameImage{images/mapk-erk.png}

% Many other biological processes can be interpreted as changes in activity

% \frame{ \frametitle{Activity-based models}
%   Many biological processes can be described as changes of activity.
% }

% \frame{ \frametitle{Timed Automata}
%   \begin{center}
%     \pgfuseimage{ta}
%   \end{center}
% }

\frame[t]{ \frametitle{Analysis of Networks with Interactive MOdelling}
\only<2-2>{\vspace{-5ex}{\begin{center}\pgfuseimage{animo-workflow}\end{center}}}%
  \begin{itemize}
    \item<1-1,3->{Interaction based }\only<1-1>{\ \\\begin{center}\pgfuseimage{animo-user-interface}\end{center}} \vspace{1ex}
    \item<3-> Discrete concentration/activity levels \vspace{1ex}
  \end{itemize}
}

\frame{ \frametitle{Discrete activity levels}
  \begin{center}
      \only<1-1>{\pgfuseimage{activity00}}\only<2-2>{\pgfuseimage{activity01}}\only<3-3>{\pgfuseimage{activity02}}\only<4-4>{\pgfuseimage{activity03}}\only<5->{\pgfuseimage{activity04}}%
  \end{center}%
  \ \\
  \uncover<6->{Let the user choose granularity: 2 \--\ 100 discrete levels}
}

%   \frame{ \frametitle{Different granularity choices (10 levels)}
%     \begin{center}
%       \pgfuseimage{grafico-ruvido}
%     \end{center}
%   }
% 
%   \frame{ \frametitle{Different granularity choices (100 levels)}
%     \begin{center}
%       \pgfuseimage{grafico-liscio}
%     \end{center}
%   }

\def\graficoStartingActivity{\includegraphics[scale=0.1]{images/starting_activity_graph}}
\newlength\graficoHeight
\setlength\graficoHeight{\heightof{\graficoStartingActivity}}
\def\scenarioDialog{\includegraphics[scale=0.27]{images/scenario_parameter_dialog}}
\newlength\dialogHeight
\setlength\dialogHeight{\heightof{\scenarioDialog}}

\frame[t]{ \frametitle{Analysis of Networks with Interactive MOdelling}
  \begin{itemize}
    \item<1-> Interaction based \vspace{1ex}
    \item<1-> Discrete concentration/activity levels \vspace{1ex}
    \item<1-> Precise reactions $\Rightarrow$ abstract \emph{interactions}\only<2-3>{$$
\mbox{\it E} + \mbox{\it S} + \mbox{\it ATP} \rightleftarrows \mbox{\it ES} + \mbox{\it ATP} \rightarrow \mbox{\it ES}^{\mbox{\scriptsize \it P}} + \mbox{\it ADP} \rightleftarrows \mbox{\it E} + \mbox{\it S}^{\mbox{\scriptsize\it P}} + \mbox{\it ADP}
$$
(with $\mbox{\it S} + \mbox{\it S}^{\mbox{\scriptsize\it P}} = \mbox{constant}$ and $\mbox{\it ATP} + \mbox{\it ADP} = \mbox{constant}$)\\
\only<3-3>{\begin{center}\Large$\Downarrow$\end{center}\vspace{-1ex}\ \\
\begin{center}\includegraphics[width=0.3\textwidth]{images/interaction_E_S}\end{center}}} \vspace{1ex}
    \item<4-> Simplified scenarios for rate computation\\\hspace{-0.8cm}\begin{minipage}[c][\dialogHeight]{0.4\textwidth}\begin{center}\scenarioDialog\end{center}\end{minipage} \begin{minipage}[c][\dialogHeight]{0.3\textwidth}\begin{center}\includegraphics[scale=0.015]{images/parameter_network}\vspace{2ex}\\\pgfuseimage{activity-legend}\end{center}\end{minipage} \begin{minipage}[c][\dialogHeight]{0.2\textwidth}\begin{center}\includegraphics[scale=0.02]{images/parameter_graph}\end{center}\end{minipage} \vspace{1ex}
%\only<5-5>{\begin{minipage}[c][\dialogHeight]{0.3\textwidth}\begin{center}\includegraphics[scale=0.015]{images/scenario1-2_network}\vspace{2ex}\\\pgfuseimage{activity-legend}\end{center}\end{minipage} \begin{minipage}[c][\dialogHeight]{0.2\textwidth}\begin{center}\includegraphics[scale=0.02]{images/scenario1-2_graph}\end{center}\end{minipage}}
    %\item<6-> Reactant activity levels determine interaction rates\only<8->{\begin{center}\begin{minipage}[c][\graficoHeight]{0.15\textwidth}\begin{center}\includegraphics[scale=0.35]{images/activity_legend_vertical}\end{center}\end{minipage}\quad\begin{minipage}[c][\graficoHeight]{0.3\textwidth}\begin{center}\includegraphics[scale=0.065]{images/starting_activity_network}\end{center}\end{minipage}\qquad\begin{minipage}[c][\graficoHeight]{0.35\textwidth}\begin{center}\graficoStartingActivity\end{center}\end{minipage}\end{center}} \vspace{1ex}
  \end{itemize}
}

\frame{ \frametitle{Timed Automata model}
  \begin{center}
    \pgfuseimage{ta-reactant-centered}
  \end{center}
}

\frame{\frametitle{ANIMO workflow}
\only<1-1>{Start from static network topology\\\begin{center}\includegraphics[width=0.9\textwidth]{images/rete_iniziale}\end{center}}
\only<2-2>{Add kinetics and choose initial activities\\\begin{center}\includegraphics[width=0.9\textwidth]{images/rete_con_valori}\end{center}}
\only<3-5>{Inspect system evolution on the fly\vspace{-1.2ex}\\\begin{center}
\begin{tikzpicture}
  \draw <3-> node (1) at (0,0) {\includegraphics[width=0.9\textwidth]{images/rete_dinamica}};
  \draw <4-> node (2) at (2.1,1.3) {\includegraphics[width=0.5\textwidth]{images/grafico1}};
  \draw <5-> node (3) at (2.1,-1.4) {\includegraphics[width=0.5\textwidth]{images/grafico2}};
\end{tikzpicture}\end{center}}
}

\frame{\frametitle{Experimental data as reference}
\begin{center}
  \includegraphics[width=0.8\textwidth]{images/chondrocyteModel}\ \\
Use data and model to improve knowledge, generate hypotheses.
\end{center}
}

\def\graphA{\includegraphics[scale=0.13]{images/feedback_params_graphA}}
\newlength\graphASize
\setlength\graphASize{\heightof{\graphA}}

\frame{\frametitle{All good and nice, but\dots}
{\Huge How to find the parameters?}\\
\begin{itemize}
  \item<1-> Insert parameters manually\only<1-1>{\\\includegraphics[scale=0.3]{images/scenario_parameter_dialog}}
  \item<2-> Compare model versions \emph{subtracting} their activity graphs%\only<2-2>{\\\begin{center}\begin{minipage}[c][\graphASize]{.8\textwidth}\begin{center}\graphA \topskip0pt\vspace*{\fill}{\Huge -}\vspace*{\fill} \includegraphics[scale=0.13]{images/feedback_params_graphD}\end{center}\end{minipage}\ \\\xdash[15em]\\ \includegraphics[scale=0.1]{images/difference_network_d-a_larger}\end{center}}
\only<2-2>{\\\begin{center}\begin{tabular}[c]{m{\graphASize}m{1em}m{\graphASize}}\graphA & \begin{center}{\Huge -}\end{center} & \includegraphics[scale=0.13]{images/feedback_params_graphD}\end{tabular}\ \\\xdash[15em]\\ \includegraphics[scale=0.1]{images/difference_network_d-a_larger}\end{center}}
   \item<3-> Perform automatic parameter scans\only<3-3>{\\\includegraphics[width=.8\textwidth]{images/parameters_results_with_bars}}
\end{itemize}

}

\frame{\frametitle{ANIMO live demo}
\begin{center}
\includegraphics[width=0.8\textwidth]{images/animo_live_demo}
\end{center}
}

% \frame{ \frametitle{Case study: \emph{in silico} experiments}
%   \begin{center}
%     %  \begin{minipage}[c]{.9\textwidth}
%           \pgfuseimage{case-study}\\[-2ex]
%     %  \end{minipage}
%   \end{center}\vspace{-1ex}
%   {\tiny Model\vspace{-1.3ex} inspired by S. D. M. Santos, P. J. Verveer, and P. I. H. Bastiaens, ``Growth factor-induced
% MAPK network topology shapes Erk response determining PC-12 cell fate'', \emph{Nature Cell Biology}, Feb. 2007.}
% }

\frame{ \frametitle{Conclusions}
  \begin{itemize}
    \item<1-> Cytoscape $\rightarrow$ static representation,\\Timed Automata $\rightarrow$ dynamic behaviour
    %\item User interface based on Cytoscape: keep the complexity of Timed Automata \emph{under the hood}
    \item<2-> Discrete levels and simplified scenarios $\rightarrow$ few parameters
    %\item Underlying Timed Automata model provides formal foundations
    \item<3-> Modelling with ANIMO:\\
      \begin{itemize}
          \item outline pathway topology
          \item select reaction parameters
          \item compare model with existing data
          \item repeat until satisfied
      \end{itemize}
    \item<4-> ANIMO allows biologists to draw network ``sketches''
%    \item<5-> Complex models:\\
%      \begin{itemize}
%	\item perform \emph{in silico} experiments
%	\item use model checking
%      \end{itemize}
     \item<5-> Parameter choice:\\
	\begin{itemize}
		\item manual settings, choice of qualitative parameters
		\item comparison of different model versions
		\item parameter sweeps
	\end{itemize}
  \end{itemize}
}

\frame[t]{ \frametitle{Future work}
  \begin{itemize}
%     \item Advance the development of user interface
%     \item Model checking queries:\begin{itemize} \item user-friendly (templates)
%     \item probabilistic (statistical model checking)
%     \end{itemize}
   \item<1-> Deal with parameter sensitivity and model \emph{robustness}\only<2-3,5->{\\Typically, biological networks are robust: smaller parameter variations do not change the behavior of the network}
    \item<3-> Get a model by feeding data: Automata Learning
    \item<4-> Use \emph{in-silico experiments} to infer biological hypotheses which can be verified through in-vitro experiments
      \only<4-4>{\vspace{1ex}\\\includegraphics[width=0.9\textwidth]{images/large_model_sox9_runx2}}
    %\item<3-> High-performance analysis of large models: multicore and distributed model checking
     % \item sensitivity analysis and parameter fitting
    %\end{itemize}
    \item<5-> Abstraction techniques to deal with large models
  \end{itemize}
}

\frame{ \frametitle{Thank you}
  \begin{center}
      \pgfuseimage{animo-logo}\\
      Analysis of Networks with Interactive MOdelling\\[2ex]
      {\Large \url{http://fmt.cs.utwente.nl/tools/animo}}\\[1ex]
      \pgfuseimage{animo-qrcode}
  \end{center}
}

\end{document}